Value-Suppressing Uncertainty Maps

### 2017-09-19 paper/main.tex
JH edits.

We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment. Our results indicate that VSUPs create close integration between uncertainty and data, resulting in better performance at data tasks with uncertainty components.

@@

\begin{enumerate}
	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $5$ ships on safe locations on the board.''
\end{enumerate}

@@

The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value (lowest danger), ignoring the uncertainty information. However, as with roulette and other similar games of chance, the \emph{variability} in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}. A risky player would choose guesses with high expected value, regardless of the uncertainty of those points. A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty. We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.

@@

We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and encourage \emph{caution} by highlighting the ambiguity introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would avoid targets with \textbf{high uncertainty} when using a VSUP.
	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
\end{enumerate}

@@

These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty. In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in lower expected values. However, when value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful. Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=0.24$) conditions.

>>>>

We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment. Our results indicate that VSUPs create close integration between uncertainty and data, resulting in a demonstrably effect of decision-making, leading viewers to make more cautious choices that avoid high uncertainty.

@@

\begin{enumerate}
	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to minimize danger. E.g., ``place your $5$ ships on the safest locations on the board.''
\end{enumerate}

@@

The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the lowest predicted danger (highest expected value), ignoring the uncertainty information. However, as with roulette and other similar games of chance, the \emph{variability} in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}. A risky player would choose guesses with high expected value, regardless of the uncertainty of those points. A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty. We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.

@@

We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and encourage \emph{caution} by highlighting the ambiguity introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would avoid targets with \textbf{high uncertainty} when using a VSUP.
	\item This would result in a tradeoff where they would also choose targets with lower expected value \textbf{higher danger} when using a VSUP.
\end{enumerate}

@@

These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty. In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in guesses with higher average danger. However, when expected value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful. Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=0.24$) conditions.
